Conditional selectivity for triplestores

### 2012-09-18 conditional/trunk/selectivity-basic.tex
inequalities subsubsection
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@381 43ef1190-90fd-4dd7-a769-d0121f02b80e

\begin{abstract}
  This paper explores selectivity estimation for single triple
  patterns in SPARQL queries. In this paper we see how the common
  assumption of independence of the selectivity of RDF terms can be
  abandoned. We proceed to extend SPLENDID\cite{splendid} using the
  new formalism and demonstrate how VoID data descriptions\cite{void} can be
  further exploited in the conditional regime. Finally, we will
  explore how selectivity can be determined exactly when a SPARQL
  query is executed on a fully indexed triplestore, such as
  Hexastore\cite{Weiss:2008:HSI:1453856.1453965}. Finally, we will
  suggest further directions for selectivity estimation in the
  conditional regime.

@@

There will be a number of different equations that arises from using
the commutative property of intersections, and some parts of these
equations is possible to compute from VoID descriptions. We have
already noted the most trivial case, for variables we have:

@@

In the case where the object is a is a variable, there may be multiple
objects for each predicate in the partition, so we can only say:
\begin{equation}\label{eq:selOvPp}
 \sel{O = ? | P = p} \ge
   \frac{\rdfterm{void:propertyPartition/void:distinctObjects}}{\rdfterm{void:propertyPartition/void:triples}} .
\end{equation}
We note that OWL cardinality constraints may constrain this problem
and that \rdfterm{void:vocabulary} may be used to find the
corresponding ontology.

@@

>>>>

\begin{abstract}
  This paper explores selectivity estimation for single triple
  patterns in SPARQL queries. In this paper we see how the common
  assumption of independence of the selectivity of RDF terms can be
  abandoned. We proceed to extend SPLENDID\cite{splendid} using the
  new formalism and demonstrate how VoID data descriptions\cite{void}
  can be further exploited in the conditional regime to optimize join
  order. Finally, we will explore how selectivity can be determined
  exactly when a SPARQL query is executed on a fully indexed
  triplestore, such as
  Hexastore\cite{Weiss:2008:HSI:1453856.1453965}. Finally, we will
  suggest further directions for selectivity estimation in the
  conditional regime.

@@

There will be a number of different equations that arises from using
the commutative property of intersections, and some parts of these
equations is possible to compute from VoID descriptions.

\subsubsection{Equalities}

We have already noted the most trivial case, for variables we have:

@@

@@

\subsubsection{Inequalities}

As noted in the abstract, the goal is to establish join order for
optimization. For this to be successful, it is not critical to be able
to estimate the selectivity in all cases. It may also suffice to
establish that the selectivity of one triple pattern is greater or
smaller than another, they can then be ordered.

First, in the case where the object is a variable, there may be multiple
objects for each predicate in the partition, so we can only say:
\begin{equation}\label{eq:selOvPp}
 \sel{O = ? | P = p} \ge
   \frac{\rdfterm{void:propertyPartition/void:distinctObjects}}{\rdfterm{void:propertyPartition/void:triples}} .
\end{equation}
We note that OWL cardinality constraints may constrain this problem
and that \rdfterm{void:vocabulary} may be used to find the
corresponding ontology.
